Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-05-07 PAPER/main.tex
added ham cycle defn, moved sub problem defn, finished draft MFFD+

\begin{figure}[h!]	
	\centering
	\includestandalone[width=0.8\textwidth]{figures/itemsknife}
	\caption{Dimensions of an item $i \in \mathcal{I}$, and knives mounted on a bar score two items simultaneously.}	
	\label{fig:itemsknife}
\end{figure}

@@

The remainder of this article will firstly, in Section 2, detail a polynomial-time algorithm that is able to produce a solution to any instance of the COP, if a solution exists. Section 3 describes three heuristics that will be used to find feasible solutions to the SCSPP, and discusses the associated advantages and disadvantages. A comparison of these heuristics and an analysis of the results will be provided in Section 4, and finally Section 5 concludes the paper and proposes some potential directions for future work.

@@

The values in $\mathcal{M}$, and thus the vertices on $G$, are arranged in pairs. This is represented on $G$ by adding a set of ``blue'' edges, $B$, that contains edges between vertices that are ``partners'', i.e. whose values make up a pair in $\mathcal{M}$. It can be seen that $B$ is a perfect matching, with $|B|= n+1$.

@@

Thus, we have an undirected graph $G$ with vertex set $V = \{v_1, ..., v_{2n+2}\}$, and two distinct edge sets $B$ and $R$ such that $B \cap R = \emptyset$. Figure \ref{fig:partners/threshold} illustrates the construction of $G$ using the instance $\mathcal{M}$ from the example above.
% Equation $v_i + v_j \geq \tau$ provided $p(v_i) \neq v_j$, i.e. not partners
% Bijective function $p: V \to V$ assoc each vertex with its partner, $v_j = p(v_i)$
% Values are assigned from $\mathcal{M}$ in non-decreasing order, i.e. value assoc with v_1 \leq value assoc with v_2 etc..

\begin{figure}	
\centering
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/partners}
	\caption{$G = (V, B)$}	
	\label{fig:partners}
\end{subfigure} \hspace{40pt}
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/threshold}
	\caption{$G=(V, B \cup R)$}	
	\label{fig:threshold}
\end{subfigure}
\caption{Example construction of $G$ using instance $\mathcal{M} = \{\{1,2\}, \{1,7\}, \{2,4\}, \{3,5\}, \{3,6\}, \{4,4\}\}$ and $\tau = 7$.}
\label{fig:partners/threshold}
\end{figure}

Consider the following definition describing a variant of the classical Hamiltonian cycle on a graph $G$.
\begin{definition}
	\label{defn:althamcycle}
	Let $G(V, B\cup R)$ be an undirected graph, where each edge is a member of one of two sets, $B$ or $R$. $G$ contains an alternating Hamiltonian cycle if there exists a Hamiltonian cycle such that successive edges alternate between sets $B$ and $R$. For example, if $(v_{i-1}, v_i) \in B$, then $(v_i, v_{i+1}) \in R$, or vice versa.
\end{definition}

@@

To find a matching $R' \subseteq R$, the following \textsc{Maximum Cardinality Matching} (MCM) algorithm can be executed, as suggested by \cite{mahadev1994} and \cite{becker2010}: For each vertex $v_i, i=1,2,...,2n+2$, select the edge in $R$ connecting $v_i$ to the highest indexed vertex $v_j$ that is not already incident to an edge in $R'$. Add this edge $(v_i, v_j)$ to $R'$, and proceed to the next vertex until all the vertices have been assessed. The two vertices incident to each matching edge in $R'$ are now referred to as being ``matched''.

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	\item $v_{i-1} \in R'$; and
	\item $(v_{i-1}, p(v_i)) \in R$.
\end{enumerate*}
Then, $v_{i-1}$ is matched with $p(v_i)$, and $v_i$ is matched with the vertex that was previously matched with $v_{i-1}$. It is guaranteed that MCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time (\citealp{mahadev1994}).

@@

\section{Solving the SCSPP}
\label{sec:scsppsoln}

\begin{definition}
	Let $\mathcal{I}' \subseteq \mathcal{I}$ be a set of items whose total width is less than or equal to the capacity of a strip, (i.e. $\sum_{i=1}^{|\mathcal{I}'|}w_i \leq W$). Given a minimum scoring distance $\tau$, the Score-Constrained Packing Sub-Problem (SCPSP) involves finding an arrangement of the items in $\mathcal{I}'$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$.
\end{definition}

@@

\subsection{Modified First-Fit Decreasing with AHCA (MFFD$^+$)}
\label{sec:mffd+}
\begin{itemize}
	\item Like original FFD, sort items in non-increasing order, and attempt to pack each item into first strip that can feasibly accommodate the item without exceeding capacity
	\item if item can fit onto strip, apply AHCA on all items on the strip and the current item to find a feasible alignment of the items
	\item if solution found, place the items on the strip in the order of the solution
	\item if no solution found, no feasible alignment exists, attempt to place current item on next strip
	\item Using AHCA means if a feasible alignment of items inc current items exists (i.e. VSC is met between all items on strip), it will be found
	\item Unlike in Basic FFD where an item cannot be placed on the end because VSC is not met, here we can find a solution
	\item items may potentially be entirely rearranged, not in non-decreasing order on each strip as in Basic FFD
	\item Slower as AHCA must be run for each item that can fit on a strip
	\item see Figure \ref{fig:comparestrips}, item cannot be placed onto strip in Basic FFD heuristic as VSC is not met in either orientation of the item, whereas by using FFD+ a feasible solution can be found, item can be placed, no need to try another strip or start a new strip
\end{itemize}

@@

>>>>

\begin{figure}[h!]	
	\centering
	\includestandalone[width=0.8\textwidth]{figures/itemsknife}
	\caption{Dimensions of an item $i \in \mathcal{I}$ in a regular orientation $(a_i, b_i)$, and a feasible alignment of two items that can be scored simultaneously.}	
	\label{fig:itemsknife}
\end{figure}

@@

\begin{definition}
	\label{defn:subprob}
	Let $\mathcal{I}' \subseteq \mathcal{I}$ be a set of items whose total width is less than or equal to the capacity of a strip, (i.e. $\sum_{i=1}^{|\mathcal{I}'|}w_i \leq W$). Given a minimum scoring distance $\tau$, the Score-Constrained Packing Sub-Problem (SCPSP) involves finding an arrangement of the items in $\mathcal{I}'$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$.
\end{definition}

The remainder of this article will firstly, in Section 2, detail a polynomial-time algorithm that is able to produce a solution to any instance of the COP, if a solution exists. Section 3 describes three heuristics that can be used to find feasible solutions to the SCSPP, and discusses the associated advantages and disadvantages. A comparison of these heuristics and an analysis of the results will be provided in Section 4, and finally Section 5 concludes the paper and proposes some potential directions for future work.

@@

The values in $\mathcal{M}$, and thus the vertices on $G$, are arranged in pairs. This is represented on $G$ by adding a set of ``blue'' edges, $B$, that contains edges between vertices that are ``partners'', i.e. whose values make up a pair in $\mathcal{M}$. It can be seen that $B$ is a perfect matching, with $|B|= n+1$. Let the bijective function $p: V \to V$ associate each vertex with its partner, $p(v_i) = v_j$. Hence $B = \{(v_i, p(v_i)) : v_i \in V\}$

@@

Thus, we have an undirected graph $G$ with vertex set $V = \{v_1, ..., v_{2n+2}\}$ and two distinct edge sets $B$ and $R$ such that $B \cap R = \emptyset$. Figure \ref{fig:partners/threshold} illustrates the construction of $G$ using the instance $\mathcal{M}$ from the example above.
% Equation $v_i + v_j \geq \tau$ provided $p(v_i) \neq v_j$, i.e. not partners
% Values are assigned from $\mathcal{M}$ in non-decreasing order, i.e. value assoc with v_1 \leq value assoc with v_2 etc..

\begin{figure}	
\centering
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/partners}
	\caption{$G = (V, B)$}	
	\label{fig:partners}
\end{subfigure} \hspace{40pt}
\begin{subfigure}[h]{0.45\textwidth}
	\includestandalone[width=\textwidth]{figures/threshold}
	\caption{$G=(V, B \cup R)$}	
	\label{fig:threshold}
\end{subfigure}
\caption{Construction of $G$ using instance $\mathcal{M} = \{\{1,2\}, \{1,7\}, \{2,4\}, \{3,5\}, \{3,6\}, \{4,4\}\}$ and $\tau = 7$, where $v_{13}$ and $v_{14}$ are the dominating vertices.}
\label{fig:partners/threshold}
\end{figure}

Recall that a Hamiltonian cycle on a graph $G$ is a cycle that visits every vertex exactly once. Now, consider the following definition describing a variant of the classical Hamiltonian cycle involving multiple edge sets.
\begin{definition}
	\label{defn:althamcycle}
	Let $G(V, B\cup R)$ be an undirected graph, where each edge is a member of one of two sets, $B$ or $R$. $G$ contains an alternating Hamiltonian cycle if there exists a Hamiltonian cycle such that successive edges alternate between sets $B$ and $R$. For example, if $(v_{i-1}, v_i) \in B$, then $(v_i, v_{i+1}) \in R$, or vice versa.
\end{definition}

@@

To find a matching $R' \subseteq R$, the following \textsc{Maximum Cardinality Matching} (MCM) algorithm can be executed, as suggested by \cite{mahadev1994} and \cite{becker2010}: For each vertex $v_i, i=1,2,...,2n+2$, select the edge in $R$ connecting $v_i$ to the highest indexed vertex $v_j$ that is not already incident to an edge in $R'$. Add this edge $(v_i, v_j)$ to $R'$, and proceed to the next vertex until all the vertices have been assessed. The two vertices incident to each matching edge in $R'$ are now referred to as being ``matched''. Similarly to partners, let $m : V \to V$ be a bijective function that assigns each vertex with its match, $m(v_i) = v_j$. Then, $R' =\{(v_i, m(v_i)) : v_i \in V\}$.

During MCM, if a vertex $v_i$ is not adjacent to any other unmatched vertex except its partner $p(v_i)$, the preceeding vertex $v_{i-1}$ can be ``rematched'', provided that 
\begin{enumerate*}[label={(\alph*)}]
	\item $v_i$ is not the first vertex; 
	%\item $v_{i-1} \in R'$; and
	\item $(v_{i-1}, m(v_{i-1})) \in R'$; and
	\item $(v_{i-1}, p(v_i)) \in R$.
\end{enumerate*}
Then, $v_i$ is matched with the vertex that is currently matched with $v_{i-1}$, i.e $m(v_i) = m(v_{i-1})$, and $v_{i-1}$ is rematched with $p(v_i)$, i.e. $m(v_{i-1}) = p(v_i)$. It is guaranteed that MCM will produce a maximum cardinality matching $R' \subseteq R$, and is able to do so in at most $O(n \log n)$ time (\citealp{mahadev1994}).

@@

\section{Heuristics for the SCSPP}
\label{sec:scsppsoln}

@@

\subsection{Modified First-Fit Decreasing with AHCA (MFFD$^+$)}
\label{sec:mffd+}

The last heuristic incorporates AHCA as described in Section \ref{sec:ahca}. The \textsc{Modified First-Fit Decreasing with AHCA} (MFFD$^+$) heuristic operates in a similar fashion to the MFFD, placing items sorted in decreasing order onto the lowest-indexed strip. However, when the lowest-indexed strip is found that can accommodate the current item, MFFD$^+$ then executes AHCA on the sub-problem, where the items on the strip and the current item form the subset $\mathcal{I}' \subseteq \mathcal{I}$ (see Definition \ref{defn:subprob}). If AHCA finds a feasible solution, the items are placed on the strip in the order of the solution, which includes the current item, else MFFD$^+$ attempts to pack the current item on the next lowest-indexed strip. Using AHCA means that if a feasible alignment of the items exists, there is a guarantee that it will be found. Unlike MFFD, where the current item can only be placed on the end of the strip, MFFD$^+$ allows the items to be entirely rearranged (see Figure \ref{fig:comparestrips}). This reduces the possibility of having to start a new strip for an item, thus preventing increasing the number of strips in the final solution.

@@

\textbf{Speed $n$ = 500}
\begin{itemize}
	\item MFFD $W = 5000$ Average speed = 38.88504ms/instance
	\item MFFD $W = 2500$ Average speed = 38.55795ms/instance
	\item MFFD $W = 1250$ Average speed = 37.98091ms/instance
	\item PS $W = 5000$ Average speed = 42.11938ms/instance
	\item PS $W = 2500$ Average speed = 44.24521ms/instance
	\item PS $W = 1250$ Average speed = 49.66614ms/instance
	\item MFFD$^+$ $W = 5000$ Average speed = 214.04479ms/instance
	\item MFFD$^+$ $W = 2500$ Average speed = 190.36579ms/instance
	\item MFFD$^+$ $W = 1250$ Average speed = 214.78729ms/instance
\end{itemize}

\textbf{Speed $n$ = 1000}
\begin{itemize}
	\item MFFD $W = 5000$ Average speed = 159.68755ms/instance
	\item MFFD $W = 2500$ Average speed = 
	\item MFFD $W = 1250$ Average speed = 
	\item PS $W = 5000$ Average speed = 
	\item PS $W = 2500$ Average speed = 
	\item PS $W = 1250$ Average speed = 
	\item MFFD$^+$ $W = 5000$ Average speed = 
	\item MFFD$^+$ $W = 2500$ Average speed = 
	\item MFFD$^+$ $W = 1250$ Average speed = 
\end{itemize}
